The K2 Bright Star Survey

### 2019-05-17 paper/ms.tex
all bedding changes made

% describe saturation
The \kepler detector saturates at a magnitude of $K_p \sim 11.3$ in both long- (30\,min) and short (1\,min)-cadence data, as these both represent sums of 6~s exposures \citep{Gilliland2010}. For objects brighter than this, additional photons create excess electrons which `bleed' into adjacent pixels in both directions along the column containing the bright star. Simple aperture photometry (SAP) -- adding all the flux contained in a window around the bleed column -- has recovered light curves with precisions close to the photon noise limit of stars as bright as 16~Cyg~AB, $\theta$~Cyg and RR~Lyr \citep[e.g.][]{Kolenberg2011,2013MNRAS.433.1262W,Guzik2016}. In the nominal \kepler mission this was not attempted for all such bright stars, and in \ktwo, the several-pixel spacecraft motion significantly increased the size of the required apertures for SAP photometry of very saturated stars, while also making their systematics more difficult to deal with. While the second-version pixel-level-decorrelation (PLD) pipeline EVEREST~2.0 was able to correct systematics in saturated SAP photometry \citep{everest2}, this is rendered impossible for the very brightest stars whose bleed columns may run to the edge of the detector. Furthermore, bandwidth constraints meant that pixel data were not downloaded for many bright targets in \ktwo.

@@

@@

% describe halo algorithm
This method was first described by \citet{White2017} and applied to the Pleiades' Seven Sisters, and was applied to Aldebaran with further developments by \citet{Farr2018}. We follow the `OWL' concept described by Hogg \&  Foreman-Mackey (2014, unpublished: preprint  \href{https://github.com/davidwhogg/OWL/}{\nolinkurl{github.com/davidwhogg/OWL/}}) in our assumptions. We assume that a star has a wide PSF sampled by many pixels with different sensitivities, which moves around on the detector within a small region and which may vary to a small extent in time. We do not have metadata describing the spacecraft motion, pixel gains, PSF variations or noise processes. Because photometry is a linear operation, any estimator of the flux is necessarily a weighted sum of pixel values, which we choose to be time-invariant but note that this strong constraint is not necessary in general.

The additional constraint beyond the OWL axioms is that some pixels are saturated, so that SAP is not possible. We consider instead the unsaturated pixels $p_j$ at the wings of the broad and structured PSF. We construct a light curve as a linear combination of these time series with weights $w_j$, so that flux at cadence $i$ is

@@

In addition to expanding the range of possible objective functions, we have also added a feature `\texttt{deathstar}' to deal with contamination. We apply the watershed-based image segmentation algorithm from \textsc{k2p2} \citep{k2p2} to the input target pixel file datacube to identify possible background sources and cut them out. \textcolor{red}{More here from Tim.} Other than this, we have adopted less-aggressive quality flagging, having found that many cadences were being classified as bad quality for spurious `cosmic ray' events, which were actually caused by a combination of saturation and spacecraft motion. We now iteratively sigma-clip outliers and retain cadences with the \texttt{lightkurve} default quality mask.

% incorporation of k2sc
While the halo procedure produces in most cases a fairly clean light curve, there are nevertheless residual systematic errors related to spacecraft motion. In order to correct these, we employ the \textsc{k2sc} code \citep{Aigrain2015,k2sc}, which simultaneously models a light curve as a 3D Gaussian Process (GP) in time and predicted position (the K2 standard data product \textsc{pos\_corr}) in pixels $(x,y)$. The model prediction in time for fixed position is then a nonparametric model of the stellar variability, and the prediction for the $x,y$ component evaluated for fixed time represents the pointing systematics. We subtract the systmatics model from the input fluxes to obtain a final corrected flux, which is the time series we use and recommend for science. The target pixel files for C91, C92, and C101 include no position information, and there are no halo apertures for C112. As a result \textsc{k2sc}-corrected data are not available for these targets.

@@

In order to choose the values for $k$ and $\delta$ in our objective function, we have selected the quiet star 36~Ophiuchi, the lowest-mass main sequence star in the sample of stars with halo apertures, and one with very little high frequency variability detected or predicted, which was also observed at short cadence. We choose the 6.5~hour Combined Differential Photometric Precision \citep[CDPP,][]{cdpp} as implemented in \texttt{lightkurve} \citep{lightkurve} as a proxy for the `noise' in a lightcurve, with lower being better. We calculate halo lightcurves and their CDPPs for $k \in \{1,2,3\}$, and $\delta \in [1,50]$ for long cadence and for various valies of $\delta \in [1,2500]$ for short cadence data. The results are displayed in Figures~\ref{cdpps_lc} and~\ref{cdpps_sc}. We find that for long cadence data, $k=1$ (TV) and a lag $\delta=10$ provide the best CDPP. This is unsurprising: that this is one cadence shorter than the 12~cadence thruster firing period. In this context we can understand the optimum as suppressing systematics on the same timescale as they occur. On the other hand, for short cadence data, performance at short lags is very poor but the method performs similarly for $k \in \{1,2\}$ with slow improvement with larger $\delta$, and performs very poorly for $k=3$ at all lags.

@@

% figure

@@

We now seek to establish the global noise properties of the whole unsaturated sample, and compare these to PDC. We process all~2466 stars with TV-min and $\delta = 10$, using all pixels in the TPF unmasked. Because these stars are so bright and the TPFs so small, in the great majority  of cases we do not expect significant contamination, and this is a way of testing how well the weights assigned by TV-min match the flux distribution over pixels. For each lightcurve we calculate the 6.5\,hr CDPP proxy with \texttt{lightkurve} as a measure of SNR, and we plot the results of the two pipelines against one another in Figure~\ref{fig:halovspdc}. We see that a significant number of stars have high PDC CDPP but low TV-min CDPP, which raises the possibility that these are variables for which halo is overcorrecting. We find by inspection of the weightmaps and \kepler pipeline aperture masks that these mostly consist of stars for which the SAP aperture is significantly smaller than the PSF. In this case, by ignoring the pipeline apertures, \texttt{halophot} is in fact generating significantly better light curves. Over all stars, we find that the fractional enclosed halo weight in the \kepler pipeline aperture is only $0.19 \pm 0.11$, which suggests that in fact the apertures are systematically smaller than optimal for stars of this magnitude, and that TV-min is using information in the fainter pixels to help correct systematics.

@@

Histograms of the CDPPs of the SAP, PDC and halo light curves with and without \textsc{k2sc} are displayed in Figure~\ref{fig:cdpphists}. We see that both halo and PDC significantly outperform SAP, with halo performing better than PDC with no additional correction. Nevertheless, after \textsc{k2sc}, we find that the best PDC lightcurves have a smaller CDPP than the best similarly pointing-corrected halo. We conjecture that PDC with its improved calibration for common-mode systematics and blended/background light is correcting for effects that halo, as a single-star and instrument-agnostic method, does not.

@@

The full sample of stars for which halo apertures were obtained is listed in Table~\ref{table_all}. While some very bright stars were observed with conventional apertures as part of these programs, simple aperture photometry is satisfactory on these targets and we exclude them from the present discussion and data release, which is oriented strictly towards targets only observable with halo photometry. 
% normal apertures in 6
We make an exception for Spica, which was observed in Campaign~6 without a halo aperture but in Campaign~17 with a halo aperture. In Campaign~6 it was assigned a normal aperture by mistake and simple aperture photometry performed extremely poorly, so we have processed it with the halo pipeline. The stars in Campaign~18 were also on-silicon in Campaign~5, but were not assigned apertures suitable for halo photometry in C5. A possible further extension of the present work would be to recover C5 light curves for these objects using either or both of smear or modified halo photometry.

% short cadence availability 
Seven stars in Campaign~13 and one in Campaign~16 were not only assigned long-cadence halo apertures, but also downloaded at short cadence. For these targets we have provided both long and short cadence reductions. Following the analysis in Section~\ref{method} showing the insensitivity of short cadence CDPP to lags longer than $\sim 100$~cad and on choice of objective function, for consistency with long cadence we have adopted a 300~cadence lag (i.e. $30 \times$ the long cadence lag of 10) and the L1 TV objective function. With their long timeseries the short cadence stars are computationally intractable for the Gaussian Process model in \textsc{k2sc} and we present otherwise uncalibrated halo lightcurves.

@@

Analyses for several of the objects here have been previously published, but we here provide the first public data releases for the Pleiades' Seven Sisters \citep{White2017}, Aldebaran \citep{Farr2018}, $\iota$~Lib \citep{Buysschaert2018}, and $\epsilon$~Tau \citep{Arentoft2019}, as well as $\rho$~Leo which was studied with halo pixels but without our objective functions \citep{Aerts2018}.

@@

% interesting EB - 98 Tau
The chemically-peculiar A0V star 98~Tau is of special interest for studies of surface inhomogeneity. We detect variability with a fundamental period of 1.74~d with twice as much power at the first harmonic ($P = 0.87$\,d), which is consistent with $\alpha^2$\,CVn spot modulation from a rapidly-rotating star with a period of 1.74\,d. This star also experiences a transit of depth 0.03, which for a 1.87\,\rsun typical A0V star imply an 0.3 \rsun companion, almost certainly of stellar mass. With rotational modulation and an eclipse to break degeneracies, models such as \texttt{starry} \citep{starry} can infer surface brightness maps and reveal the spatial distribution of the star's chemical peculiarity.

@@

% do hot star asteroseismology and red giants as benchmarks, interferometry
Some of the objects presented here are the subject of more detailed work in preparation, namely Spica (Buzasi et al., in prep.), interferometry of the Hyades giants (White et al., in prep.), and main-sequence stars (Greklek-McKeon et al., in prep.).

% do k2 smear after that
The sample of K2 bright stars presented here only includes those with halo apertures, but while some others are available conventionally, many were not assigned target pixels and were not downloaded at all. Smear photometry has been used to recover the brightest otherwise-unobserved stars in nominal \kepler (Pope et al., under review), and this can also be done in K2, although in the latter case the sample is much smaller due to competition with halo apertures and the systematics correction is more challenging. A natural extension of both pieces of work would be to produce smear light curves of all bright stars without halo apertures in K2, which would finally make the \kepler extended mission magnitude-complete at the bright end.

% tess halo 
The halo method naturally extends to other contexts where simple aperture photometry is not possible, such as for saturated stars observed by the Transiting Exoplanet Survey Satellite \citep[TESS;][]{tess}. Although the saturation limit is brighter ($T_{mag} \sim 6$) and this problem accordingly affects fewer stars and less badly, there are situations such as for $\alpha$~Centauri or $\beta$~Hydri where the bleed column reaches the edge of the chip and a SAP light curve is irrecoverable. We expect that TV-min halo photometry will therefore be valuable in ensuring that TESS can observe even the very brightest stars.

>>>>

% describe saturation
The \kepler detector saturates at a magnitude of $K_p \sim 11.3$ in both long- (30\,min) and short (1\,min)-cadence data, since these both represent sums of 6\,s exposures \citep{Gilliland2010}. For objects brighter than this, excess electrons `bleed' into adjacent pixels in both directions along the column containing the star. Simple aperture photometry (SAP) -- adding all the flux contained in a window around the bleed column -- has recovered light curves with precisions close to the photon noise limit of the $V = 6$ 16~Cyg~AB, $V=4.48$ $\theta$~Cyg, and the $V=7.2$ RR~Lyr \citep[e.g.][]{Kolenberg2011,2013MNRAS.433.1262W,Guzik2016}. In the nominal \kepler mission this was only attempted for a few bright stars, and in \ktwo, the larger spacecraft motion significantly increased the size of the required apertures for SAP photometry of very saturated stars, while also making their systematics more difficult to deal with. While the second-version pixel-level-decorrelation (PLD) pipeline EVEREST~2.0 was able to correct systematics in saturated SAP photometry \citep{everest2}, this is not possible for the very brightest stars whose bleed columns may run to the edge of the detector. Furthermore, bandwidth constraints meant that pixel data were not downloaded for many bright targets in \ktwo.

@@

\newpage

@@

% describe halo algorithm
This method was first described by \citet{White2017} and applied to the Pleiades' Seven Sisters. It was also applied to Aldebaran with further developments by \citet{Farr2018}. We follow the `OWL' concept described by Hogg \&  Foreman-Mackey (2014, unpublished: preprint  \href{https://github.com/davidwhogg/OWL/}{\nolinkurl{github.com/davidwhogg/OWL/}}) in our assumptions. We assume that a star has a wide PSF sampled by many pixels with different sensitivities, which moves around on the detector within a small region and which may vary to a small extent in time. We do not have metadata describing the spacecraft motion, pixel gains, PSF variations or noise processes. Because photometry is a linear operation, any estimator of the flux is necessarily a weighted sum of pixel values, which we choose to be time-invariant but note that this strong constraint is not necessary in general.

The additional constraint beyond the OWL axioms is that some pixels are saturated, so that SAP is not possible. Instead the measurements are made using the unsaturated pixels $p_j$ at the wings of the broad and structured PSF. We construct a light curve as a linear combination of these time series with weights $w_j$, so that flux at cadence $i$ is

@@

In addition to expanding the range of possible objective functions, we have also added a feature `\texttt{deathstar}' to deal with contamination. We apply the watershed-based image segmentation algorithm from \textsc{k2p2} \citep{k2p2} to the input target pixel file datacube to identify possible background sources and cut them out. \textcolor{red}{More here from Tim.} Other than this, we have adopted less-aggressive quality flagging, having found that many cadences were being classified as bad quality for spurious `cosmic ray' events, which were actually caused by a combination of saturation and spacecraft motion. We instead chose to iteratively sigma-clip outliers and retain cadences with the \texttt{lightkurve} default quality mask.

% incorporation of k2sc
While the halo procedure produced a fairly clean light curve in most cases, there were nevertheless residual systematic errors related to spacecraft motion. In order to correct these, we employed the \textsc{k2sc} code \citep{Aigrain2015,k2sc}, which simultaneously models a light curve as a 3D Gaussian Process (GP) in time and predicted position (the K2 standard data product \textsc{pos\_corr}) in pixels $(x,y)$. The model prediction in time for fixed position is then a nonparametric model of the stellar variability, and the prediction for the $x,y$ component evaluated for fixed time represents the pointing systematics. We subtracted the systmatics model from the input fluxes to obtain a final corrected flux, which is the time series we use and recommend for science. Campaigns~9, 10, and~11 were observed in two blocks each, denoted C91/C92, C101/C102 and C111/C112 by the \ktwo Team. The target pixel files for C91, C92, and C101 include no position information, and there are no halo apertures for C112. As a result \textsc{k2sc}-corrected data are not available for these targets.

@@

In order to choose the values for $k$ and $\delta$ in our objective function, we used the quiet star 36~Ophiuchi, the lowest-mass main sequence star in the sample of stars with halo apertures, and one with very little high frequency variability detected or predicted. It was also observed at short cadence. We chose the 6.5~hour Combined Differential Photometric Precision \citep[CDPP,][]{cdpp} as implemented in \texttt{lightkurve} \citep{lightkurve} as a proxy for the `noise' in a lightcurve, with lower being better.

We calculated halo lightcurves of 36\,Oph and their CDPPs for $k \in \{1,2,3\}$, and $\delta \in [1,50]$ for long cadence and for various valies of $\delta \in [1,2500]$ for short cadence data. The results are displayed in Figures~\ref{cdpps_lc} and~\ref{cdpps_sc}. We found that for long cadence data, $k=1$ (TV) and a lag $\delta=10$ provide the best CDPP. This is unsurprising: that this is one cadence shorter than the 12~cadence thruster firing period. In this context we can understand the optimum as suppressing systematics on the same timescale as they occur. On the other hand, for short cadence data, performance at short lags is very poor but the method performs similarly for $k \in \{1,2\}$ with slow improvement with larger $\delta$, and performs very poorly for $k=3$ at all lags.

@@

@@

We now seek to establish the global noise properties of the whole unsaturated sample, and compare these to PDC. We process all~2466 stars with TV-min and $\delta = 10$, using all pixels in the TPF unmasked. Because these stars are so bright and the TPFs so small, in the great majority  of cases we do not expect significant contamination, and this is a way of testing how well the weights assigned by TV-min match the flux distribution over pixels. For each lightcurve we calculated the 6.5\,hr CDPP proxy with \texttt{lightkurve} as a measure of SNR, and we plot the results of the two pipelines against one another in Figure~\ref{fig:halovspdc}. We see that a significant number of stars have high PDC CDPP but low TV-min CDPP, which raises the possibility that these are variables for which halo is overcorrecting. We found by inspection of the weightmaps and \kepler pipeline aperture masks that these mostly consist of stars for which the SAP aperture is significantly smaller than the PSF. In this case, by ignoring the pipeline apertures, \texttt{halophot} is in fact generating significantly better light curves. Over all stars, we found that the fractional enclosed halo weight in the \kepler pipeline aperture is only $0.19 \pm 0.11$, which suggests that in fact the apertures are systematically smaller than optimal for stars of this magnitude, and that TV-min is using information in the fainter pixels to help correct systematics.

@@

Histograms of the CDPPs of the SAP, PDC and halo light curves with and without \textsc{k2sc} are displayed in Figure~\ref{fig:cdpphists}. We see that both halo and PDC significantly outperform SAP, with halo performing better than PDC with no additional correction. Nevertheless, after \textsc{k2sc}, we found that the best PDC lightcurves have a smaller CDPP than the best similarly pointing-corrected halo. We conjecture that PDC with its improved calibration for common-mode systematics and blended/background light is correcting for effects that halo, as a single-star and instrument-agnostic method, does not.

@@

The full sample of stars for which halo apertures were obtained is listed in Table~\ref{table_all}. Some very bright stars were observed with conventional apertures as part of these programs, but we exclude them from the present discussion and data release, which is oriented towards targets only observable with halo photometry. 
% normal apertures in 6
We include $\alpha$\,Vir (Spica), which was observed in Campaign~6 without a halo aperture but in Campaign~17 with a halo aperture. In Campaign~6 it was assigned a normal aperture by mistake and simple aperture photometry performed extremely poorly, so we have processed these data with the halo pipeline. The stars in Campaign~18 in our sample were also on-silicon in Campaign~5, but were not assigned apertures suitable for halo photometry in C5. A possible further extension of the present work would be to recover C5 light curves for these objects using smear and/or modified halo photometry.

% short cadence availability 
Seven stars in Campaign~13 and one in Campaign~16 were assigned short-cadence halo apertures. For these targets we have provided both long- and short-cadence reductions. Following the analysis in Section~\ref{method} showing the insensitivity of short cadence CDPP to lags longer than $\sim 100$\,cad and on choice of objective function, and for consistency with long cadence, we have adopted a 300~cadence lag (i.e. $30 \times$ the long cadence lag of 10) and the L1 TV objective function. With their many time samples, the short-cadence stars are computationally intractable for the Gaussian Process model in \textsc{k2sc} and we present otherwise uncalibrated halo lightcurves.

@@

Analyses for some of our sample have been previously published, we include their light curves in this data release: the Pleiades Seven Sisters \citep{White2017}, $\alpha$~Tau \citep[Aldebaran;][]{Farr2018}, $\iota$~Lib \citep{Buysschaert2018}, and $\epsilon$~Tau \citep{Arentoft2019}, as well as $\rho$~Leo, which was studied with halo pixels but without our objective functions \citep{Aerts2018}.

@@

% interesting EB - 98 Tau
The chemically-peculiar A0V star 98~Tau is of special interest for studies of surface inhomogeneity. We detected variability with a fundamental period of 1.74~d with twice as much power at the first harmonic ($P = 0.87$\,d), which is consistent with $\alpha^2$\,CVn spot modulation from a rapidly-rotating star with a period of 1.74\,d. This star also experiences a transit of depth 0.03, which for a 1.87\,\rsun typical A0V star implies an 0.3\,\rsun stellar mass companion. With rotational modulation and an eclipse to break degeneracies, models such as \texttt{starry} \citep{starry} can infer surface brightness maps and reveal the spatial distribution of the star's chemical peculiarity.

@@

% do hot star asteroseismology and red giants as benchmarks, interferometry
Some of the objects presented here are the subject of more detailed work in preparation, namely $\alpha$\,Vir (Spica), interferometry of the Hyades giants, and main-sequence stars.

% do k2 smear after that
The sample of K2 bright stars presented here only includes those with halo apertures. While some others are available conventionally, many were not assigned target pixels and were not downloaded at all. Smear photometry has been used to recover the brightest otherwise-unobserved stars in nominal \kepler (Pope et al., under review), and this can also be done in K2, although the sample is much smaller due to allocation of halo apertures and the systematics correction is more challenging. A natural extension of both pieces of work would be to produce smear light curves of all bright stars without halo apertures in K2, which would finally make the \kepler extended mission magnitude-complete at the bright end.

% tess halo 
The halo method naturally extends to other contexts where simple aperture photometry is not possible, such as for saturated stars observed by the Transiting Exoplanet Survey Satellite \citep[TESS;][]{tess}. Although the saturation limit is brighter ($T_{mag} \sim 6$) and this problem accordingly affects fewer stars and less badly, there are stars such as $\alpha$~Centauri and $\beta$~Hydri where the bleed column reaches the edge of the chip and a SAP light curve is irrecoverable. We expect that TV-min halo photometry will therefore be valuable in ensuring that TESS can observe the very brightest stars.

% halo method
